---
title: "HW2_Markdown"
output: github_document
author: Noah Kreski
---
``` {r setup, include = FALSE, warning = FALSE}
library(tidyverse)
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)


```
#Problem One
```{r question_one_dataset}
subway_data = read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")%>%
              janitor::clean_names(dat = .)%>%
              select(.data = ., line:entry, vending, ada)
              subway_data$entry <- recode(subway_data$entry,"'YES' = TRUE;'NO'=FALSE")
              subway_data$entry <- as.logical(subway_data$entry)
               
              
              

```


#Problem Two



#Problem Three

```{r question_three_dataset}

#Formatting appropriate names
brfss_smart2010_Tidy = janitor::clean_names(brfss_smart2010)%>%
                      #Removing unneeded variables
                      select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location))%>%
                      #Isolating necessary data values
                      filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"))%>%
                      #Splitting by response category
                      spread(key=response, value = data_value)%>%
                      janitor::clean_names()%>%
                      mutate(EVG = excellent + very_good)

brfss_smart2010_Tidy_2002 = filter(brfss_smart2010_Tidy, year == 2002)
```

###How many unique locations are included in the dataset? 

There are `r nrow(distinct(brfss_smart2010_Tidy, locationdesc))` locations, based on the distinct number of location descriptions.

###Is every state represented? 

Every state is represented given that there are `r nrow(distinct(brfss_smart2010_Tidy, locationabbr))` location abbreviations which account for all states and Washington D.C.


###What state is observed the most?
```{r state_freq}

table(brfss_smart2010_Tidy$locationabbr)

```

Using the table function, which listed the number of observations for each location abbreviation, the 50 states and Washington DC, it can be seen that New Jersey, with 146 observations, is the most observed.


###In 2002, what is the median of the “Excellent” response value?

The median of the Excellent response value in 2002 is `r median(brfss_smart2010_Tidy_2002$excellent, na.rm=TRUE)` using the median function.


###Make a histogram of “Excellent” response values in the year 2002.

```{r excellent_historgram, warning = FALSE}
ggplot(brfss_smart2010_Tidy_2002, aes(x = excellent)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

###Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.
```{r data_NY, include = FALSE}
brfss_smart2010_Tidy_NY = filter(brfss_smart2010_Tidy, locationdesc %in% c("NY - New York County","NY - Queens County"))
```
```{r scatterplot_NY}
ggplot(brfss_smart2010_Tidy_NY, aes(x = year, y = excellent)) + 
  geom_point(aes(color = locationdesc))
```
