---
title: "HW2_Markdown"
output: github_document
author: Noah Kreski
---

```{r question_three_dataset}
# install.packages("devtools")
devtools::install_github("p8105/p8105.datasets")

library(p8105.datasets)
#Formatting appropriate names
brfss_smart2010_Tidy = janitor::clean_names(brfss_smart2010) %>%
                      #Removing unneeded variables
                      select(-class, -topic, -question, -sample_size, -(confidence_limit_low:geo_location))%>%
                      #Isolating necessary data values
                      filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"))%>%
                      #Splitting by response category
                      spread(key=response, value = data_value)%>%
                      janitor::clean_names()%>%
                      mutate(EVG = excellent + very_good)
```

How many unique locations are included in the dataset? Is every state represented? What state is observed the most?

In 2002, what is the median of the “Excellent” response value?

```{r excellent_median,  include=FALSE}
brfss_smart2010_Tidy_2002 = filter(brfss_smart2010_Tidy, year == 2002)
```

The median of the Excellent response value in 2002 is ```r median(brfss_smart2010_Tidy_2002$EVG)```.

Make a histogram of “Excellent” response values in the year 2002.

Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.

